Dominance Regions for Affine Cluster Algebras

### 2022-06-16 affine_dominance.tex
Section 4
Computed the c-vector part of the extended exchange matrix for all preprojective seeds

\newcommand{\bfc}{\boldsymbol{c}}
\newcommand{\bfg}{\boldsymbol{g}}
\newcommand{\bfr}{\boldsymbol{r}}
\newcommand{\bfx}{\boldsymbol{x}}
\newcommand{\bfy}{\boldsymbol{y}}

@@

  The index $k\in\{1,\ldots,n\}$ also determines a new matrix $\mu_k \tilde{B}=(b'_{ij})$ given by
  \[ b'_{ij}=\begin{cases} -b_{ij} & \text{if $i=k$ or $j=k$;}\\ b_{ij}+[b_{ik}]_+b_{kj}+b_{ik}[-b_{kj}]_+ & \text{otherwise.} \end{cases} \]
    One easily observes that the principal part of $\mu_k \tilde{B}$ is again skew-symmetrizable using the same matrix $D$.
  \begin{remark}
    Note that $\mu_k \tilde{B}=E_{k,\varepsilon} \tilde{B} F_{k,\varepsilon}$ for $\varepsilon=\pm 1$, the case $\varepsilon=1$ being obvious from the definitions and the case $\varepsilon=-1$ following from the identity $b_{ij}=[b_{ij}]_+-[-b_{ij}]_+$.
  \end{remark}

@@

  Suppose the principal part $B$ of $\tilde{B}$ is acyclic.
  Given a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, define a quiver $Q_{B,\varepsilon}$ from $Q_B$ by removing all arrows starting at vertex $i$ whenever $\varepsilon_i=-1$.
  For $1\le j\le n$, define $\cP_j(Q_{B,\varepsilon})$ as the set of paths in $Q_{B,\varepsilon}$ starting at vertex $j$.
  Elements of $\cP_j(Q_{B,\varepsilon})$ are sequences of vertices $(p_1,\ldots,p_r)$ with $p_1=j$ and $b_{p_{t+1}p_t}>0$ for $1\le t\le r-1$.
  Given a path $p\in\cP_j(Q_{B,\varepsilon})$, a vertex $i\in Q_0$, and a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, set
  \[\wt_{i,\varepsilon}(p):=-b_{ip_r}^{(1-\delta_{ip_r})(1+\varepsilon_{p_r})/2}\prod_{t=1}^{r-1} b_{p_{t+1}p_t},\]
  where the prefactor $b_{ip_r}^{(1-\delta_{ip_r})(1+\varepsilon_{p_r})/2}$ is omitted if $i=p_r$ or $\varepsilon_{p_r}=-1$.
  \begin{lemma}
    Suppose $B$ is acyclic with source-adapted sequence $k_1,\ldots,k_n$.
    Consider $t_1,\ldots,t_n\in\TT_n$ so that $t_{\ell-1}$ is joined to $t_\ell$ by an edge in $\TT_n$ labeled by $k_\ell$ for $1\le\ell\le n$.
    For a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, the product $E^{t_{n-1}}_{k_n,\varepsilon_{k_n}}\cdots E^{t_0}_{k_1,\varepsilon_{k_1}}$ is given by $\tau=(\tau_{ij})$ with
    \begin{equation}
      \label{eq:tropical twist}
      \tau_{ij}=\sum_{p\in\cP_j(Q_{B,\varepsilon})}^* \wt_{i,\varepsilon}(p),
    \end{equation}
    where the $*$ indicates that the summation includes the trivial path at $j$ only if $i\le j$ when $\varepsilon_j=1$ and only if $i=j$ when $\varepsilon_j=-1$.

@@

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A$ be an invertible $n\times n$ matrix.
    Under the left action of $A$, say $A\bfx=\bfy$ for $\bfx,\bfy\in\RR^n$, the region inside $\RR^n$ defined by the inequality $\langle\bfr,\bfx-\lambda\rangle\le t$ with $\lambda,\bfr\in\RR^n$ and $t\in\RR$ is transformed into the region defined by the inequality \[\langle A^{-T}\bfr,\bfy-A\lambda\rangle\le t.\]
  \end{lemma}
  \begin{proof}
    This is immediate from the equalities
    \[\langle\bfr,\bfx-\lambda\rangle=\bfr^T(\bfx-\lambda)=\bfr^TA^{-1}A(\bfx-\lambda)=(A^{-T}\bfr)^T(A\bfx-A\lambda)=\langle A^{-T}\bfr,\bfy-A\lambda\rangle.\]
  \end{proof}

@@

>>>>

\newcommand{\bfc}{\boldsymbol{c}}
\newcommand{\bfe}{\boldsymbol{e}}
\newcommand{\bfg}{\boldsymbol{g}}
\newcommand{\bfn}{\boldsymbol{n}}
\newcommand{\bfr}{\boldsymbol{r}}
\newcommand{\bfu}{\boldsymbol{u}}
\newcommand{\bfx}{\boldsymbol{x}}
\newcommand{\bfy}{\boldsymbol{y}}

@@

  The index $k\in\{1,\ldots,n\}$ also determines a new matrix $\mu_k \tilde{B}=(b'_{ij})$ given by
  \begin{equation}
    \label{eq:matrix mutation}
    b'_{ij}=\begin{cases} -b_{ij} & \text{if $i=k$ or $j=k$;}\\ b_{ij}+[b_{ik}]_+b_{kj}+b_{ik}[-b_{kj}]_+ & \text{otherwise.} \end{cases}
  \end{equation}
    One easily observes that the principal part of $\mu_k \tilde{B}$ is again skew-symmetrizable using the same matrix $D$.
  \begin{remark}
    Note that $\mu_k \tilde{B}=E_{k,\varepsilon} \tilde{B} F_{k,\varepsilon}$ for $\varepsilon=\pm 1$, the case $\varepsilon=1$ being obvious from the definitions and the case $\varepsilon=-1$ following from the identity $b_{ij}=[b_{ij}]_+-[-b_{ij}]_+$.
  \end{remark}

@@

@@

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A$ be an invertible $n\times n$ matrix.
    Under the left action of $A$, say $A\bfx=\bfy$ for $\bfx,\bfy\in\RR^n$, the region inside $\RR^n$ defined by the inequality $\langle\bfn,\bfx-\lambda\rangle\le t$ with $\lambda,\bfn\in\RR^n$ and $t\in\RR$ is transformed into the region defined by the inequality \[\langle A^{-T}\bfn,\bfy-A\lambda\rangle\le t.\]
  \end{lemma}
  \begin{proof}
    This is immediate from the equalities
    \[\langle\bfn,\bfx-\lambda\rangle=\bfn^T(\bfx-\lambda)=\bfn^TA^{-1}A(\bfx-\lambda)=(A^{-T}\bfn)^T(A\bfx-A\lambda)=\langle A^{-T}\bfn,\bfy-A\lambda\rangle.\]
  \end{proof}

@@

  \section{Attempts at understanding $\tau$}

  Suppose the principal part $B$ of $\tilde{B}$ is acyclic.
  Given a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, define a quiver $Q_{B,\varepsilon}$ from $Q_B$ by removing all arrows starting at vertex $i$ whenever $\varepsilon_i=-1$.
  For $1\le j\le n$, define $\cP_j(Q_{B,\varepsilon})$ as the set of paths in $Q_{B,\varepsilon}$ starting at vertex $j$.
  Elements of $\cP_j(Q_{B,\varepsilon})$ are sequences of vertices $(p_1,\ldots,p_r)$ with $p_1=j$ and $b_{p_{t+1}p_t}>0$ for $1\le t\le r-1$.
  Given a path $p\in\cP_j(Q_{B,\varepsilon})$, a vertex $i\in Q_0$, and a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, set
  \[\wt_{i,\varepsilon}(p):=-b_{ip_r}^{(1-\delta_{ip_r})(1+\varepsilon_{p_r})/2}\prod_{t=1}^{r-1} b_{p_{t+1}p_t},\]
  where the prefactor of $b_{ip_r}$ is omitted if $i=p_r$ or $\varepsilon_{p_r}=-1$.
  \begin{lemma}
    Suppose $B$ is acyclic with source-adapted sequence $k_1,\ldots,k_n$.
    Consider $t_1,\ldots,t_n\in\TT_n$ so that $t_{\ell-1}$ is joined to $t_\ell$ by an edge in $\TT_n$ labeled by $k_\ell$ for $1\le\ell\le n$.
    For a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, the product $E^{t_{n-1}}_{k_n,\varepsilon_{k_n}}\cdots E^{t_0}_{k_1,\varepsilon_{k_1}}$ is given by $\tau=(\tau_{ij})$ with
    \begin{equation}
      \label{eq:tropical twist}
      \tau_{ij}=\sum_{p\in\cP_j(Q_{B,\varepsilon})}^* \wt_{i,\varepsilon}(p),
    \end{equation}
    where the $*$ indicates that the summation includes the trivial path at $j$ only if $i\le j$ when $\varepsilon_j=1$ and only if $i=j$ when $\varepsilon_j=-1$.

\section{Iterated Mutations}

  Let $B=(b_{ij})$ denote an acyclic $n\times n$ exchange matrix.
  Write $Q_B$ for the corresponding quiver, here $b_{ij}>0$ means there is an arrow from $j$ to $i$ in $Q$.
  Without loss of generality we assume $1,\ldots,n$ is a sink adapted sequence for $Q_B$, i.e. vertex $1$ is a sink and vertex $n$ is a source in $Q$.
  This implies $B$ has positive entries above the diagonal.
  \sayDR{This seemed to give the cleanest calculation.}
  Define vectors $\bfc^{(k)}_r\in\ZZ^n$ for $1\le k\le n$ and $r\in\ZZ_{\ge0}$ recursively as follows:
  \sayDR{A slight modification is needed to handle the opposite mutation sequence.}
  \begin{itemize}
    \item $\bfc^{(k)}_0=\bfe_k$ is the standard basis vector 
    \item for $1\le \ell \le n$, we have $\bfc^{(k)}_{ns+\ell}=\begin{cases} -\bfc^{(k)}_{ns+\ell-1} & \text{if $\ell=k$;}\\ \bfc^{(k)}_{ns+\ell-1}+b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1} & \text{if $\ell<k$;}\\ \bfc^{(k)}_{ns+\ell-1}-b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1} & \text{if $\ell>k$.} \end{cases}$
  \end{itemize}
  For $r\in\ZZ$ and $1\le k\le n$, write $r\equiv_n k$ if $n|(r-k)$ and $r\not\equiv_n k$ otherwise.
  \begin{lemma}
    \label{lem:tropical signs}
    Assume $B$ is not finite-type.
    For $r\in\ZZ_{\ge0}$, the vector $\bfc^{(k)}_r$ is non-negative if $r\not\equiv_n k$ and non-positive when $r\equiv_n k$.
  \end{lemma}
  \begin{proof}
    Induction?
  \end{proof}

  \begin{lemma}
    For any $s\in\ZZ_{\ge0}$ and $0\le \ell \le n-1$, the matrix $\mu_\ell\cdots\mu_1(\mu_n\cdots\mu_1)^s\tilde B$ has principal part given by $(b'_{ij})$ where $b'_{ij}=(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij}$ for $\epsilon(i,j)=\begin{cases} 1 & \text{if $i\le j$;}\\ 0 & \text{otherwise;}\end{cases}$ and the $\bfc$-vector part of the $k$-th column is given by $\bfc^{(k)}_{ns+\ell}$ for $1\le k\le n$.
  \end{lemma}
  \begin{proof}
    We work by induction on $ns+\ell$.
    For $s=\ell=0$, we just get $\tilde B$ and the claim clearly holds.
    Assume the claim holds for a chosen $s$ and $\ell-1$, we show the claim holds for $s$ and $\ell$ (the case $\ell=n$ gives rise to the claim for $s+1$ and $\ell=0$).
    Write $\tilde B'=\mu_{\ell-1}\cdots\mu_1(\mu_n\cdots\mu_1)^s\tilde B$ and let $Q'_B$ denote the corresponding acyclic quiver.

    By \eqref{eq:matrix mutation}, the mutation $\mu_\ell$ reverses the sign of the $\ell$-th row and column of $\tilde B'$.
    It follows that the principal part of $\mu_\ell\tilde B'$ is given as follows:
    \begin{itemize}
      \item if $i,j<\ell$, or $i<\ell$ and $j>\ell$, or $i<\ell$ and $j>\ell$, then $(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij}=(-1)^{\epsilon(i,\ell-1)+\epsilon(j,\ell-1)}b_{ij}$;
      \item if $i=\ell$ or $j=\ell$, then $(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij}=-(-1)^{\epsilon(i,\ell-1)+\epsilon(j,\ell-1)}b_{ij}$;
    \end{itemize}
    in each case we obtain the desired expression.
    In addition, we see that the $\ell$-th $\bfc$-vector is given by $-\bfc^{(\ell)}_{ns+\ell-1}=\bfc^{(\ell)}_{ns+\ell}$ as desired.
    By Lemma~\ref{lem:tropical signs}, the $\ell$-th $\bfc$-vector of $\tilde B'$ is non-negative.  
    Observe further that vertex $\ell$ is a sink in $Q'_B$ so that the $\ell$-th row of $\tilde B'$ has non-negative entries, in particular we see that $b'_{ij}=-b_{ij}$ for $j<\ell$ and $b'_{ij}=b_{ij}$ for $j>\ell$.
    It then follows from \eqref{eq:matrix mutation} that the $k$-th $\bfc$-vector of $\mu_\ell\tilde B'$ is given by $\bfc^{(k)}_{ns+\ell-1}+b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1}$ if $\ell<k$ and by $\bfc^{(k)}_{ns+\ell-1}-b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1}$ if $\ell>k$.
    Each of these is equal to $\bfc^{(k)}_{ns+\ell}$ as desired.
  \end{proof}
